### 2021-02-14 rad-icfp-2021/macrodiff.tex
initial commit ifcp paper

\section{A macro for source-to-source reverse-mode differentiation}