### 2021-02-14 rad-icfp-2021/macrodiff.tex
some comments

\section{A macro for source-to-source reverse-mode differentiation}

\MH{for forward it's as in the fossacs paper. for rad it's a bit more tricky.}