### 2021-02-16 rad-icfp-2021/simplecase.tex
some very rough starting draft

\section{Simple case}

\subsection{Language}

We consider a simple first-order language. 
TODO

\subsection{Naive reverse-mode}

\subsection{Optimisations}

\subsection{Correctness}

\subsection{Cheap gradient principle}

\subsubsection*{Key insight:}

As is well known (see e.g. \cite{pearlmutter2008reverse}), duplication of a variable $\Delta$ is turned into $+$ by reverse mode.
TODO: explain this is the hard part to make pure and efficient, and shows up in the case of non-unary operators, and things like let binding.
In the inefficient representation, every operator was seen as a unary operator using all the available variables and returning them in addition to the extra return value. 
In essence, we are putting boxes locally around places where the compute-flow might be non-linear and turning everything into something linear.
This linearisation of the compute-flow is the key to efficient purely functional reverse-mode. We extend this idea formally in the next sections by introducing a unary form, 
on which it is easier to do naive reverse-mode that can be optimized.