### 2021-02-22 rad-icfp-2021/efficientdiff.tex
change outline

\section{Efficient reverse-mode}

\subsection{Target Language}

\subsection{Efficient representation}

\subsection{Optimisations via partial evaluation}