### 2021-02-22 rad-icfp-2021/simplediff.tex

\section{Simple reverse-mode differentiation}

\subsection{Source Language}

\subsection{Source Unary Form}

\subsection{Target Unary Form}

\subsection{Macro for reverse mode}