### 2021-02-22 rad-icfp-2021/efficientdiff.tex
adding some structure to the document

\section{Efficient reverse-mode}

\subsection{Target Language}

\input{figures/target_grammar}

\input{figures/target_type_system}

\input{figures/unf_to_target}

\subsection{Efficient representation}

\subsection{Optimisations via partial evaluation}