### 2021-02-22 rad-icfp-2021/simplediff.tex

\section{Simple reverse-mode differentiation}

\subsection{Source Language}

\input{figures/source_grammar}

\input{figures/source_type_system}

\subsection{Source Unary Form}

\input{figures/source_unf}

\input{figures/source_to_unf}

\subsection{Target Unary Form}

\input{figures/target_unf}

\subsection{Macro for reverse mode}

\input{figures/macro_diff}