### 2021-02-23 rad-icfp-2021/figures/compilation_scheme.tex
adding some figures

\[
\begin{tikzcd}
    \Source \ar[r] & \UNFSymbol_1 \ar[r,"\Dsynrevsymbol"] & \UNFSymbol_2 \ar[r] & \Target \ar[r,"\partial ev"] & \Target
\end{tikzcd}
\]