### 2021-02-23 rad-icfp-2021/generalisation.tex
minor

\section{Generalisations}

Extended source language:

\input{figures/generalisation_source_grammar}

\input{figures/generalisation_source_typesystem}

Extended target language:

\input{figures/generalisation_target_grammar}

\input{figures/generalisation_target_typesystem}

\subsection{Arrays}

\subsection{Conditionals}

\subsection{Partial operations}

\subsection{Higher-order functions}