### 2021-02-24 rad-icfp-2021/generalisation.tex
small fixes

\section{Generalisations}
\label{sec:generalisation}

Extended source language:

\input{figures/generalisation_source_grammar}

% \input{figures/generalisation_source_typesystem}

Note that we recover !zip A B! as !map2 (fun x,y -> (x,y)) A B)!. 
We can lift the restriction on the function of !foldl! by changing a part in $\invUNFSymbol$ of !foldl!.
\begin{center}
\begin{tabular}{r c l}
    !Y((x1,...,xm),...)! & -> &  !Y((x1,...,xm) + map2! \\
        && !(fun z1 z2 -> let E = map2 (fun x,y -> z2(x,y)) A e3 in! \\
        &&  !let F = zip B E in!\\
        && !foldl (fun (x,y), acc  -> x*acc+y) z1 F)! \\
        && $\nabla_{\Gamma}$!e2! $\nabla_{\Gamma}$!e1,...)!
\end{tabular}
\end{center}

Extended target language:

% \input{figures/generalisation_target_grammar}

% \input{figures/generalisation_target_typesystem}

\subsection{Arrays}

\subsection{Conditionals}

\subsection{Partial operations}

\subsection{Higher-order functions}