### 2021-02-26 rad-icfp-2021/efficientdiff.tex
writing about target language

\section{Efficient reverse-mode}
\label{sec:efficientrad}

\subsection{Target Language}

The last language we present, which we call the target language, is an extension to the source language. 
The types and terms are presented in Figure~\ref{fig:target_grammar}.
The language contains lambda abstractions for dealing with continuations. 
We are not interested here with partial applications and our language is pure, so functions take $n$ arguments.
Lambda abstractions and applications are going to be removed during partial evaluation and this notation makes reading slightly easier.

\input{figures/target_grammar}

In addition, we require the set of operations to be closed under partial differentiation. 
In more detail, for every unary scalar operation !op1!, 
we assumed given an operator $\partial_1$!op1! whose semantics should be the derivative of !op1!.
For every binary operator !op2!, we assume given operators $\partial_1$!op2!, $\partial_2$!op2! 
representing respectively the first and second partial derivative of !op2!.
Similarly, we need to add some functions on arrays. 
!scanl! is a !foldl! that stores all intermediate results. 
!scanr! is similar, but reads the array from right to left, and also stores from right to left.
!reduce! is a !foldl! for which the function is associative. 
This means in practice that it can be computed faster.
We also include a !map3! operator. 
Though not necessary as it can be simulated with two !map2!, it is a convenient notation and can somtimes run faster in practice.
Finally, we add a somewhat strange operator which we call !shift1!. 
It takes an array of size $n$,  shifts all the element of the array by one, 
forgetting the last element, and puts a one in the first place. 
This operator naturally shows up when differentiating !foldl!.

The type system for the extended grammar of target is in Figure~\ref{fig:target_typesystem}.

\input{figures/target_type_system}

A standard semantics of this language is given in Section~\ref{sec:correctness}.

We now present our transformation from target UNF to target. 
The transformation is presented in Figure~\ref{fig:unf_to_target}.

\input{figures/unf_to_target}

\begin{remark}
    One could rightfully wonder: where did reverse-mode really happened here?
    In a way, source UNF does not seem to do much. 
    Then $\Dsynrevsymbol$ is very simple on UNF.
    So perhaps everything happens when going from target UNF to target?
    \MH{finish or remove.}
\end{remark}

\subsection{Efficient representation}

\subsection{Optimisations via partial evaluation}