### 2021-07-01 rad-icfp-2021/figures/compilation_scheme.tex
minor corrections intro

\[
\begin{tikzcd}
    \Source \ar[rrr,"\text{efficient }\Dsynrevsymbol (Fig.~\ref{fig:direct_diff_macro})"] \ar[dr,"(Fig.~\ref{fig:source_to_unf})"'] &&& \Target \arrow[r,loop right,"\text{optim }(Fig.~\ref{fig:optim})"] \\
    & Source\UNFSymbol \ar[r,"\Dsynrevsymbol (Fig.~\ref{fig:diff_macro})"] & Target\UNFSymbol \ar[ur,"(Fig.~\ref{fig:unf_to_target})"'] & 
\end{tikzcd}
\]