### 2021-07-02 rad-icfp-2021/figures/macro_diff.tex
changing UNF to make it correct

\begin{figure*}[t]
    \begin{tabular}{r c l}
        $\Dsynrevsymbol_{\rho}$(!A1,...,An!) &=& !A1,...,An,(A1x...xAn)->!$\rho$
    \end{tabular}

    \medskip
    \begin{tabular}{r c l}
    $\Dsynrevsymbol_{\rho}$(!var!$_{T;i}$) &=& !<proj!$_{T;\underline{T}->\rho;[]} \comp$ !var!$_{T,i}$!, proj!$_{[];T;\underline{T}->\rho} \icomp$ (!proj!$_{T;\underline{T}->\rho;[]}$ $\comp$ !Jvar!$_{T;i}$)!>! \\
    $\Dsynrevsymbol_{\rho}$(!op!$_{\Gamma;n}$) &=& !<op!$_{\Gamma,n}$!, Jop!$_{\Gamma,n}$!>! \\ 
    $\Dsynrevsymbol_{\rho}$(!pair!$_{T;A\times B}$) &=& \\
    $\Dsynrevsymbol_{\rho}$(!proj!$_{T1;T2;T3}$) &=& \\
    $\Dsynrevsymbol_{\rho}$(!e1!$\comp$!e2!) &=& $\Dsynrevsymbol$(!e1!); $\Dsynrevsymbol$(!e2!)\\ 
    $\Dsynrevsymbol_{\rho}$(!e1!$\pcomp$!e2!) &=& \\
    $\Dsynrevsymbol_{\rho}$(!map2!$_{T;x,y.e}$) &=& !<map2!$_{x,y.e}$!, J!$^T$!map2!$_{x,y.e}$!>! \\ 
    $\Dsynrevsymbol_{\rho}$(!reduce!$_{T;x,y.e1;e2}$) &=& !<reduce!$_{x,y.e}$!, J!$^T$!reduce!$_{x,y.e}$!>! \\  
    \end{tabular}
    \caption{Reverse-mode differentiation from Source UNF to Target UNF}
    \label{fig:diff_macro}    
\end{figure*}