### 2021-07-03 rad-icfp-2021/generalisation.tex
writing a bit generalisation section

\section{Generalisations}
\label{sec:generalisation}

\subsection{Lifting the restriction on map and fold}
\label{sub:Lifting the restriction on map and fold}

TODO: not sure about this, and my derivative for reduce/foldl is currently a bit wrong

Note that we recover !zip A B! as !map2 (fun x,y -> (x,y)) A B)!. 
We can lift the restriction on the function of !foldl! by changing a part in $\invUNFSymbol$ of !foldl!.
\begin{center}
\begin{tabular}{r c l}
    !Y((x1,...,xm),...)! & -> &  !Y((x1,...,xm) + map2! \\
        && !(fun z1 z2 -> let E = map2 (fun x,y -> z2(x,y)) A e3 in! \\
        &&  !let F = zip B E in!\\
        && !foldl (fun (x,y), acc  -> x*acc+y) z1 F)! \\
        && $\nabla_{\Gamma}$!e2! $\nabla_{\Gamma}$!e1,...)!
\end{tabular}
\end{center}

\subsection{Adding more array operators}
\label{sub:Adding more array operators}

\subsection{Adding more scalar operators} % (fold)
\label{sub:Adding more scalar operators}

We assumed the unary and binary operators were denoted by smooth functions $\RR^n\to\RR$. 
There is no additional difficulty in considering operators which are partial functions 
like division or operators which are not smooth at a point like square root.

These functions are then given intentional derivatives which provide valid derivatives 
on the domain of definition and differentiability of the operator. 
These functions are well known to be the bete noire of AD \cite{griewank2008evaluating} 
and we do not provide novel solutions to these.  
Several recent work have shown how to give semantics to such operators in the context of AD \cite{vakar2020denotational,mazza2021automatic,sherman2021,lee2020correctness}.

\subsection{Adding conditionals} % (fold)
\label{sub:Adding conditionals}

Adding conditionals to the source language can be done easily, for instance via defining

\begin{tabular}{r c l}
$\directD{\rho}{\Gamma}{Y}(\Gamma \vdash$ !if e1 then e2 else e3!) &=& !if e1 then !$\directD{\rho}{\Gamma}{Y}$!(e2) else !$\directD{\rho}{\Gamma}{Y}$!(e3)! 
\end{tabular}

The problem is that this could break the complexity of reverse mode because of the non linear usage of $Y$, and makes everything harder to optimise.

A slighly better option would be to define

\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}(\Gamma \vdash$ !if e1 then e2 else e3!) 
    &=& !let b=e1 in!   \\
    && !<if b then e2 else e3, fun (x1,...,xn,z) ->! \\
    && !Y(b*!$\directD{\rho}{\Gamma}{Y}$!(e2)(x1,...,xn,z)+!\\
    && \quad!(1-b)*!$\directD{\rho}{\Gamma}{Y}$!(e3)(x1,...,xn,z))>!
\end{tabular}

It is slighly better because both derivatives of !e2! and !e3! are put together which might unlock some optimisations, but there is still a non linear usage of the variables.
One solution, if we know we want to compute the whole gradient of the expression, is to define instead

\begin{tabular}{r c l}
    $\directD{\rho}{\Gamma}{Y}(\Gamma \vdash$ !if e1 then e2 else e3!) 
    &=& !let b=e1 in!   \\
    && !<if b then e2 else e3, fun (x1,...,xn,z) ->! \\
    && !Y((x1,...,xn)!$\widehat{+}$($\grad_\Gamma$!(e2)*b!$\widehat{+}\grad_\Gamma$!(e3)*(1-b))*z)>!\\
\end{tabular}

This time, there is linear usage of !Y,z! and the !xi!.

TODO: explain why it's correct?\\
TODO: adding conditionals does not break differentiability as long as no non smooth function like max(0,-):R->B\\
TODO: re explain how it's usually done by partial eval of the conditional first.

\subsection{Adding general arrays} % (fold)
\label{sub:Adding general arrays}

We now show how to consider arrays over any ground type $G$.

TODO: introduce sum and tensor of semantics of ground type\\
TODO: then * becomes a dot product\\
TODO: minor changes from there\\
TODO: except I'm not sure with reduce and shift1...

Though this has the right complexity, it is an interesting avenue for future research to find
better representations to allow for more optimisations. 
In particular, representations looking like Einsum \cite{van2011numpy} could be of interest 
and has been recently studied in the context of AD \cite{laue2018computing,laue2020simple}.

TODO: more generally about tensors \cite{liao2019differentiable,bernstein2020differentiating}